---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![R-CMD-check](https://github.com/hjmbigdatabowl/bdb2021/workflows/R-CMD-check/badge.svg)](https://github.com/hjmbigdatabowl/bdb2021/actions)
<!-- badges: end -->

## NFL Big Data Bowl 2021

This is our NFL Big Data Bowl 2021 submission.

#### Authors

* Hugh McCreery
* John Edwards
* Matt Kaye
* Owen McGrattan

You can find our notebook submission on [our pkgdown site](https://hjmbigdatabowl.github.io/bdb2021/articles/submission.html). The notebook contains an explanation of our methodology and results. We also made a [Shiny app](https://bdb-2021.herokuapp.com/), which lets you explore our models, results, and player ratings.

## Setup

First, install the package from Github either by cloning the repo or with `devtools` (below)

``` r
# install.packages("devtools")
devtools::install_github("hjmbigdatabowl/bdb2021", dependencies = TRUE)
```

To run the code, you'll need to have joined the [Big Data Bowl 2021 Kaggle competition](https://www.kaggle.com/c/nfl-big-data-bowl-2021/).

Once you've joined, you can get the data from [the competition's data page](https://www.kaggle.com/c/nfl-big-data-bowl-2021/data). In addition, you'll need to get the targets data, which can be found [here](https://www.kaggle.com/tombliss/additional-data-targeted-receiver/data). Once you've downloaded the data, you need to save it in `bdb2021/inst/data/` and unzip the zip archive (if you downloaded the data as a zip).

Once that's all set up, you should be good to go! You can see the [function reference page](https://hjmbigdatabowl.github.io/bdb2021/reference/index.html) for documentation on the functions, or you can run `?bdb2021::FUNCTION_NAME`. 

```{r load_the_package}
library(bdb2021)

?get_weather
```

## Code

All of our code for our submission can be found in our Github repo, [hjmbigdatabowl/bdb2021](https://github.com/hjmbigdatabowl/bdb2021).


The code for the Shiny app can be found in another repo in Github, [hjmbigdatabowl/bdb2021-shiny](https://github.com/hjmbigdatabowl/bdb2021-shiny).


